---
title: "Ghod Sub-basin CFRA Forest Selection"
author: "Elisa Stefaniak"
date: "2025-02-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file(),
#                      knitr.digits.signif = TRUE)

options(stringsAsFactors = FALSE)
library(raster)
library(tidyverse)
# library(inborutils)
library(sf)
library(terra)
library(gtools)
library(ggpubr)

set.seed(0110)
```

## The Problem

In this document I describe the process of generating afforestation maps for the Bhima Upper Basin Ghod Sub-basin. 
First we examine the Ghod Sub-basin and it's current afforestation. 

Load Ghod Sub-basin region

```{r ghod_subbasin_mask}
data_gpkg_region <- sf::read_sf("../data/maps/region.gpkg")
data_gpkg_region_2d <- st_zm(data_gpkg_region$geom)

ggplot() + 
  geom_sf(data = data_gpkg_region_2d, fill = "lightgrey") + 
  # geom_sf(data = data_gpkg_2d, fill = "grey") +
  # geom_sf(data = data_gpkg_intersected_2d, fill = "forestgreen") + 
  # geom_sf(data = data_gpkg_intersected_out_2d, fill = "lightgreen") + 
  scale_x_continuous(limits = c(73.5, 74.05)) + 
  scale_y_continuous(limits = c(18.95, 19.25))
```

Total size of the Ghod Sub-basin region is `r sum(st_area(data_gpkg_region_2d))` m2.

Get current aforestation (raster), subset to just forests and convert to shapes. This is done in a separate code file (`code/create_current_forest_polygons.R`) as the process is computing-intensive. 


```{r print_land_use}
useclass_gpkg <- rast("../data/maps/land_use_classes.tif")

## Forest area is under 0 so we make everything else NA
current_forest2 <- sf::read_sf("../out/current_forest.shp")

ggplot() +
  geom_sf(data = data_gpkg_region_2d, fill = "lightgrey") + 
  geom_sf(data = current_forest2, fill = "forestgreen", lwd = 0) + 
  # geom_tile(data = clipped_raster_df, aes(x = x, y = y, fill = "landsurface/land_use_classes")) +
  scale_x_continuous(limits = c(73.5, 74.05)) + 
  scale_y_continuous(limits = c(18.95, 19.25)) +
  guides(fill="none")
```

Current afforestation is `r sum(st_area(current_forest2))` m2 which is: `r 100 * sum(st_area(current_forest2)) / st_area(data_gpkg_region_2d)`% of the total area of the subbasin.

Now, I examine the aforestation from the CFRA regions (dark grey) and subset to only the Ghod Subbasin (light grey). 

```{r}
data_gpkg <- sf::read_sf("../data/maps/junnar_potential_CFR.gpkg")
data_gpkg_2d <- st_zm(data_gpkg$geom)

# Extract forests only in modelled area

intersected_CFRS <- st_intersects(data_gpkg_region_2d, data_gpkg_2d)

data_gpkg_intersected_2d <- data_gpkg_2d[intersected_CFRS[[1]]]
data_gpkg_intersected_out_2d <- st_intersection(data_gpkg_region_2d, data_gpkg_2d)

ggplot() +
  geom_sf(data = data_gpkg_region_2d, fill = "lightgrey") +
  geom_sf(data = data_gpkg_2d, fill = "grey") +
  # geom_sf(data = data_gpkg_intersected_2d, fill = "forestgreen") +
  geom_sf(data = data_gpkg_intersected_out_2d, fill = "lightgreen") +
  scale_x_continuous(limits = c(73.5, 74.05)) +
  scale_y_continuous(limits = c(18.95, 19.25))
```

The total potential new aforested area is `r sum(st_area(data_gpkg_intersected_out_2d))` m2 which is: `r 100 * (sum(st_area(data_gpkg_intersected_out_2d))) / st_area(data_gpkg_region_2d)`% of the total area of the subbasin. The maximum aforestation we would get in the basin is therefore: `r 100 * (sum(st_area(data_gpkg_intersected_out_2d)) + sum(st_area(current_forest2))) / st_area(data_gpkg_region_2d)`%.

If we include now also the original forests we get:

```{r}
ggplot() +
  geom_sf(data = data_gpkg_region_2d, fill = "lightgrey") +
  geom_sf(data = data_gpkg_2d, fill = "grey") +
  geom_sf(data = current_forest2, fill = "forestgreen", lwd = 0) + 
  # geom_sf(data = data_gpkg_intersected_2d, fill = "forestgreen") +
  geom_sf(data = data_gpkg_intersected_out_2d, fill = "lightgreen") +
  scale_x_continuous(limits = c(73.5, 74.05)) +
  scale_y_continuous(limits = c(18.95, 19.25))
```

## Are there intersections in the current and potential new forest areas?

## What are the CFRA forests currently?

## Random subsetting of CFRA forests

We want to look at 0-100%. Below I create a code to get 50% of the plots. 

```{r}
forest_areas <- st_area(data_gpkg_intersected_out_2d)
total_area = sum(forest_areas)

area_percentage = 0.5
total_area_aim = total_area * area_percentage

forest_areas_perm_ind <- permute(1:length(forest_areas))
forest_areas_perm <- forest_areas[forest_areas_perm_ind]
forest_areas_perm_cumsum <- cumsum(forest_areas_perm)
ind_out <- forest_areas_perm_ind[forest_areas_perm_cumsum <= total_area_aim]

data_gpkg_intersected_out_2d_subset <- data_gpkg_intersected_out_2d[ind_out]

sum(st_area(data_gpkg_intersected_out_2d_subset)) / 10000

ggplot() + 
  geom_sf(data = data_gpkg_region_2d, fill = "lightgrey") + 
  geom_sf(data = data_gpkg_2d, fill = "lightblue") +
  geom_sf(data = data_gpkg_intersected_2d, fill = "forestgreen") + 
  geom_sf(data = data_gpkg_intersected_out_2d, fill = "lightgreen") + 
  geom_sf(data = data_gpkg_intersected_out_2d_subset, fill = "pink") + 
  scale_x_continuous(limits = c(73.5, 74.05)) + 
  scale_y_continuous(limits = c(18.95, 19.25))
```

I repeat this for 20, 40, 60 and 80% and save the outputs (executed offline).

```{r eval=FALSE, echo= TRUE}

area_percentage_aim = c(0.2, 0.4, 0.6, 0.8, 1.0)
names_file = c("02", "04", "06", "08", "1")
total_area_aim = total_area * area_percentage_aim

lapply(1:5, function(x){
  ind_out <- forest_areas_perm_ind[forest_areas_perm_cumsum <= total_area_aim[x]]
  data_gpkg_intersected_out_2d_subset <- data_gpkg_intersected_out_2d[ind_out]
  st_write(data_gpkg_intersected_out_2d_subset, paste0(c("../out/forest_", names_file[x], ".gpkg"), collapse = ""))
})
  
```

Plotting the potential areas for each afforestation scenario. The pink areas correspond to the areas selected for afforestation, and the bright green are the remainder of the potential areas for reforestation. The dark greay are the current forest areas. 

```{r, fig.height= 7.08, fig.width=14.87}
data_gpkg_forest_02 <- sf::read_sf("../out/forest_02.gpkg")
forest_02 <- st_zm(data_gpkg_forest_02$geom)

data_gpkg_forest_04 <- sf::read_sf("../out/forest_04.gpkg")
forest_04 <- st_zm(data_gpkg_forest_04$geom)

data_gpkg_forest_06 <- sf::read_sf("../out/forest_06.gpkg")
forest_06 <- st_zm(data_gpkg_forest_06$geom)

data_gpkg_forest_08 <- sf::read_sf("../out/forest_08.gpkg")
forest_08 <- st_zm(data_gpkg_forest_08$geom)

data_gpkg_forest_1 <- sf::read_sf("../out/forest_1.gpkg")
forest_1 <- st_zm(data_gpkg_forest_1$geom)


plot_02 <- ggplot() +
  geom_sf(data = data_gpkg_region_2d, fill = "lightgrey") +
  geom_sf(data = current_forest2, fill = "forestgreen", alpha = 0) +
  geom_sf(data = data_gpkg_intersected_out_2d, fill= "lightgreen") + 
  geom_sf(data = forest_02, fill = "pink") +
  scale_x_continuous(limits = c(73.5, 74.05)) + 
  scale_y_continuous(limits = c(18.95, 19.25)) +
  labs(title = "20 %", hjust = 0.5)+
  theme(plot.title = element_text(hjust = 0.5))

plot_04 <- ggplot() +
  geom_sf(data = data_gpkg_region_2d, fill = "lightgrey") +
  geom_sf(data = current_forest2, fill = "forestgreen", alpha = 0) +
  geom_sf(data = data_gpkg_intersected_out_2d, fill= "lightgreen") + 
  geom_sf(data = forest_04, fill = "pink") +
  scale_x_continuous(limits = c(73.5, 74.05)) + 
  scale_y_continuous(limits = c(18.95, 19.25)) +
  labs(title = "40 %", hjust = 0.5)+
  theme(plot.title = element_text(hjust = 0.5))

plot_06 <- ggplot() +
  geom_sf(data = data_gpkg_region_2d, fill = "lightgrey") +
  geom_sf(data = current_forest2, fill = "forestgreen", alpha = 0) +
  geom_sf(data = data_gpkg_intersected_out_2d, fill= "lightgreen") + 
  geom_sf(data = forest_06, fill = "pink") +
  scale_x_continuous(limits = c(73.5, 74.05)) + 
  scale_y_continuous(limits = c(18.95, 19.25)) +
  labs(title = "60 %", hjust = 0.5)+
  theme(plot.title = element_text(hjust = 0.5))

plot_08 <- ggplot() +
  geom_sf(data = data_gpkg_region_2d, fill = "lightgrey") +
  geom_sf(data = current_forest2, fill = "forestgreen", alpha = 0) +
  geom_sf(data = data_gpkg_intersected_out_2d, fill= "lightgreen") + 
  geom_sf(data = forest_08, fill = "pink") +
  scale_x_continuous(limits = c(73.5, 74.05)) + 
  scale_y_continuous(limits = c(18.95, 19.25)) +
  labs(title = "80 %", hjust = 0.5)+
  theme(plot.title = element_text(hjust = 0.5))

plot_1 <- ggplot() +
  geom_sf(data = data_gpkg_region_2d, fill = "lightgrey") +
  geom_sf(data = current_forest2, fill = "forestgreen", alpha = 0) +
  geom_sf(data = data_gpkg_intersected_out_2d, fill= "lightgreen") + 
  geom_sf(data = forest_1, fill = "pink") +
  scale_x_continuous(limits = c(73.5, 74.05)) + 
  scale_y_continuous(limits = c(18.95, 19.25)) +
  labs(title = "100 %", hjust = 0.5)+
  theme(plot.title = element_text(hjust = 0.5))


# afforestation_plots <- ggarrange(plot_02, plot_04, plot_06, plot_08, plot_1, 
#                                        ncol = 3, nrow = 2)


plot_02
plot_04
plot_06
plot_08
plot_1
```


## Prioritised subsetting of CFRA forests

For each potential area for afforestation, we calculate the distance to the closest current forest. 
```{r}

# # for each potential area, calculate the distance to the closest current forest
# m_to_forest0 <- lapply(1:length(data_gpkg_intersected_out_2d), function(x){
#       distance_to_forest <- st_distance(data_gpkg_intersected_out_2d[x], st_union(current_forest2))
#     })
# saveRDS(m_to_forest0, "../out/m_to_forest0.rds")

# read RDS (output from code above)
m_to_forest0 <- readRDS("../out/m_to_forest0.rds")

#save only the distance to nearest forest
m_to_forest <- lapply(1:length(data_gpkg_intersected_out_2d), function(x){
  m_to_forest0[[x]][1]
})

# turn into a data frame
m_to_forest_df0 <- do.call(rbind.data.frame, m_to_forest)
colnames(m_to_forest_df0) <- "distance_m"

# add an index based on order of distance

m_to_forest_df <- m_to_forest_df0 %>% 
  mutate(position = row_number(-distance_m))

table(m_to_forest_df$distance_m)
```

Above, we see that `r table(m_to_forest_df$distance_m)[1]` out of `r length(data_gpkg_intersected_out_2d)` polygons of potential afforestation locations are 0 meters away from current forest cells, and the maximum distance is around 550 meters away. We have thus decided to randomly assign polygons for the afforestation scenarios. 

```{r}
# st_write(data_gpkg_region, "../data/maps/region.shp")
```

